Testing and Isolation Efficacy: Insights from a Simple Epidemic Model

### 2021-04-23 SIR_manuscript.tex
more updates on method sec

\begin{table}[htp]
\centering
\begin{tabular}{|c|L{2in}cc|} \hline
  Symbol & Description & Unit & Value \\ \hline
  $N_0$     & Total population size & people & $10^6$ \\ \hline
  $\omega$  & Rate of test return, i.e., rate of onward flow from ``waiting'' to ``confirmed'' or ``untested'' compartments  & 1/day & - \\ \hline
  $\gamma$ & Recovery rate & 1/day & 1/3 \\ \hline 
  $\rho$   & \percap testing intensity & 1/day & - \\ \hline 
  $\theta\_w$ & Isolation efficacy in reduction of the probability of transmission for ``waiting'' individuals & - & - \\ \hline
  $\theta\_c$ & Isolation efficacy in reduction of the probability of transmission for ``confirmed positive'' individuals & - & -  \\ \hline
  $\beta$ & Transmission rate & 1/day & 0.39 \\ \hline
  $\Lambda$ & Force of infection & 1/day & - \\ \hline
  $p_S$ & Probability of positive tests for $S$ ($= 1-\textrm{specificity}$) & - & 0 \\ \hline
  $p_I$ & Probability of positive tests for $I$ ($= \textrm{sensitivity}$) & - & 1 \\ \hline
  $p_R$ & Probability of positive tests for $R$ ($= 1-\textrm{specificity}$) & - & 0.5 \\ \hline
  $w_S, w_I, w_R$ & Relative testing weights & - &
  \begin{minipage}[t]{0.21\columnwidth}%
    Random: $\{1,1,1\}$ \\ Targeted: $\{0.3,1,1\}$
  \end{minipage} \\
  \hline
  \end{tabular}
\caption{\label{tab:params} Parameters of the model \eqref{model}.}
\end{table}

@@

The analytical calculation of the next-generation matrix and the derivation of $\Rnum$ was carried out in Maple \citep{maple14}. 
We used R \citep{r} for simulations and for plotting $\Rnum$ as a function of the underlying parameters. In particular, the contours of $\Delta$ \eqref{eq:Delta1} were plotted as in a range of a selected set of parameters of interest, $1/\omega$, $\rho$, $\theta\_w$ and $\theta\_c$, with the other parameters values kept fixed and presented in Table~\ref{tab:params}.

and isolation efficacy parameters, $\theta\_w$ and $\theta\_c$ vary from 0 to 1.
Results are shown in Figures \ref{pan} and \ref{pan2}, where  panel (a) represents the random testing ($w_S=w_I=w_R=1$), and panel (b) is representing non-random testing ($w_S=0.3$ and $w_I=w_R=1$). 
Fig. \ref{pan} reflects the changes in $\Rnum$  when the \percap testing rate is small relative to the population size, $\rho \in [0,0.01]$, and the test return rate $\omega\in [0.1,2]$.  
Fig. \ref{pan2} reflects the changes in $\Rnum$  when the \percap testing rate is large relative to the population size, $\rho \in [0,0.5]$, and the test return rate $\omega\in (0.5,2]$. 
Note that the critical contour of $\Delta=1-\frac{\gamma}{\beta}$, corresponding to $\Rnum=1$, is plotted in solid line in the Figures.

>>>>

\begin{table}[htp]
\centering
\begin{tabular}{|c|L{2in}cc|} \hline
  Symbol & Description & Unit & Value \\ \hline
  $N_0$     & Total population size & people & $10^6$ \\ \hline
  $\omega$  & Rate of test return, i.e., rate of onward flow from ``waiting'' to ``confirmed'' or ``untested'' compartments  & 1/day & - \\ \hline
  $\gamma$ & Recovery rate & 1/day & 1/3 \\ \hline 
  $\rho$   & \percap testing intensity & 1/day & - \\ \hline 
  $\theta\_w$ & Isolation efficacy in reduction of the probability of transmission for ``waiting'' individuals & - & - \\ \hline
  $\theta\_c$ & Isolation efficacy in reduction of the probability of transmission for ``confirmed positive'' individuals & - & -  \\ \hline
  $\beta$ & Transmission rate & 1/day & 0.339 \\ \hline
  $\Lambda$ & Force of infection & 1/day & - \\ \hline
  $p_S$ & Probability of positive tests for $S$ ($= 1-\textrm{specificity}$) & - & 0 \\ \hline
  $p_I$ & Probability of positive tests for $I$ ($= \textrm{sensitivity}$) & - & 1 \\ \hline
  $p_R$ & Probability of positive tests for $R$ ($= 1-\textrm{specificity}$) & - & 0.5 \\ \hline
  $w_S, w_I, w_R$ & Relative testing weights & - &
  \begin{minipage}[t]{0.21\columnwidth}%
    Random: $\{1,1,1\}$ \\ Targeted: $\{0.3,1,1\}$
  \end{minipage} \\
  \hline
  \end{tabular}
\caption{\label{tab:params} Parameters of the model \eqref{model}.}
\end{table}

@@

The analytical calculation of the next-generation matrix and simplifying the expression of $\Rnum$ was carried out in Maple \citep{maple14}. 
We used R \citep{r} for simulations, in particular,  for plotting the contours of $\Delta$ \eqref{eq:Delta1} over a range of selected set of parameters of interest, i.e., parameters that could be manipulated by public-health policy. These parameters include the isolation efficacy parameters, $\theta\_c$ and $\theta\_w$, \percap testing intensity, $\rho$, and the rate of test return, $\omega$.
The rest of model parameters kept fixed at sensible values (see Table~\ref{tab:params} for the parameter values). Note that all plots are illustrated in the scale of the mean test return time $1/\omega$ (day).  
Results are shown in Fig.\,\ref{pan} and Fig.\,\ref{pan2} with two panels; panel (a) represents the random testing, thus $w_S=w_I=w_R=1$, and panel (b) represents targeted testing, thus $w_S=0.3$ and $w_I=w_R=1$. 
To illustrate the changes in $\Rnum$ with respect to \percap testing intensity$\rho$, two sets of plots are presented. 
Fig.\,\ref{pan} reflects the changes in $\Rnum$ when $\rho$, is small relative to the population size, specifically, $\rho \in [0,0.01]$, and the test return rate $\omega\in [0.1,2]$. This is a more realistic scenario of testing as we have observed in \covid pandemic. That is, 1\% of a population are tested per day at maximum capacity, so in a population of size $N_0=10^6$, 10000 individuals will be tased per day at maximum.
In order to illustrate the non-monotonic changes in $\Rnum$ with respect to $\rho$,
Fig. \ref{pan2} with maximum capacity of $\rho$ is larger relative to the population size, $\rho \in [0,0.5]$, and the test return rate $\omega\in (0.5,2]$. 
In both Fig.\,\ref{pan} and Fig.\,\ref{pan2}, the isolation efficacy parameters, $\theta\_w$ and $\theta\_c$ vary from 0 to 1.
Note that the critical contour of $\Delta=1-\frac{\gamma}{\beta}$, corresponding to the threshold $\Rnum=1$, is plotted in solid line in the panels. Thus, a combination of $\rho$ and $\omega$ above this critical contour results in reducing $\Rnum$. Also realize that $\Rnum=\frac{\beta}{\gamma}$, about 1.017 here, for the classical SIR model which can be acheived in the context of our model when $\rho=0$, no testing, or $\theta\_w=\theta\_w=0$, no isolation.
